---
title: "Getting started with bayr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting started with bayr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
## The examples need rstanarm (Suggests); skip them if it is not installed.
knitr::opts_chunk$set(eval = requireNamespace("rstanarm", quietly = TRUE))
```

`bayr` turns the results of Bayesian regression models into tidy tables that
plug directly into `knitr`/Markdown reports. This vignette follows the
"fitting regression models" section of *New Statistics for Design Researchers*:
we simulate a small experiment, estimate two models and report their
parameters.

## Setting up

```{r libraries}
library(rstanarm)
library(bayr)
```

Load `bayr` *after* `rstanarm` (or `brms`): `bayr` registers `predict()` and
`coef()` methods for the model classes of those packages, and the package that
is loaded last provides the methods (see the README for details).

## A simulated experiment

```{r data}
set.seed(42)
N_Obs <- 20
Exp <- tibble::tibble(
  Obs = seq_len(N_Obs),
  Condition = rep(c("Experimental", "Control"), N_Obs / 2),
  age = runif(N_Obs, 18, 35),
  mu = 200 + (Condition == "Control") * 50 + age * 1,
  outcome = rnorm(N_Obs, mu, 10)
)
Exp <- as_tbl_obs(Exp)
Exp
```

`as_tbl_obs()` tags the data frame as a table of observations. Its print method
shows a handful of randomly drawn rows with a table caption instead of dumping
all observations into the report.

## Fitting the models

```{r fit}
M_1 <- stan_glm(outcome ~ Condition, data = Exp,
                chains = 2, iter = 500, seed = 42, refresh = 0)
M_2 <- stan_glm(outcome ~ Condition + age, data = Exp,
                chains = 2, iter = 500, seed = 42, refresh = 0)
```

## Parameter estimates

```{r clu-1}
clu(M_1)
```

```{r clu-2}
clu(M_2)
```

`clu()` returns a `tbl_clu`: a tibble with the posterior median (`center`) and
the credibility limits (`lower`, `upper`) for every parameter. It prints as a
formatted table and can still be processed with the usual tidy tools.

The population-level effects alone:

```{r fixef}
fixef(M_2)
```

## Inline reporting

For numbers in running text, `md_coef()` formats a single estimate with its
interval:

```{r md-coef}
C_2 <- clu(M_2)
cat(md_coef(C_2, fixef == "age", interval = TRUE))
```

The result is a plain markdown string, so it can be embedded directly in
running text.

## The posterior in long format

The underlying draws are available as a tidy posterior, one row per draw:

```{r posterior}
P_2 <- posterior(M_2)
P_2
```

`posterior()` returns a `tbl_post`, so it can be filtered and summarized like
any other tibble:

```{r posterior-filter}
dplyr::filter(P_2, type == "fixef")
```

## Predictions

```{r predict}
predict(M_2)
```

`predict()` returns a `tbl_predicted`: one row per observation, with center and
credibility limits.

## Where to go next

* `?posterior`, `?clu`, `?coef` and `?md_coef` — the extraction and reporting
  functions.
* `?predict.tbl_post_pred` — posterior predictive draws and their summaries.
* `?IC` and `?compare_IC` — information criteria for model comparison.
